---
title: "Quantitative Data Analysis"
author: "Wenzhuo Niu"
date: "2020/1/1"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

## summary
The aim of this project is to determine the relationship between WHP and persenal conditions and sociometric factors.

## preparation
```{r packages, message = FALSE}

library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(hexbin)
library(ggExtra)
library(corrplot)
library(lmtest)
library(sandwich)
library(car)

```

## analysis

### data
The herein used data can be found in the [Health Survey for England, 2016;HSE](http://doi.org/10.5255/UKDA-SN-8334-2) from UK Data Service.


```{r loaddata}
survey2016 <- read_tsv("D:/program in glasgow/quantitative data analysis/project/Health Survey for England/UKDA-8334-tab/tab/hse2016_eul.tab" )

survey2016 <- survey2016 %>% 
  select(whval, a30wk06, mins10sptb,ad10Strength,wkhrstot,dnoft3,totalwu,Sex,Age35g,topqual3 ,nssec8,eqv5,cigst2,Nactygr,porfv15)%>%
  filter(whval > 0 & a30wk06 >= 0 & mins10sptb >= 0 & ad10Strength >= 0 & wkhrstot >= 0 & wkhrstot <= 1440 & dnoft3 > 0 & totalwu > 0 & topqual3>0 & nssec8<90 & eqv5 > 0 & Age35g >= 7 & cigst2 > 0 & Nactygr >= 0, porfv15 >= 0)%>%
  drop_na() %>%
  rename(WaistHipratio = whval, walking = a30wk06, sport = mins10sptb, exertion = ad10Strength, sedentary = wkhrstot, drinkfrequency = dnoft3, drinktotal = totalwu, education = topqual3, employment = nssec8, income = eqv5, age = Age35g, cigarette = cigst2, gambling = Nactygr, fruitandveg = porfv15)
```

### preprocessing

```{r whval}
#whval:Valid Mean Waist/Hip ratio
survey2016%>%
  filter(WaistHipratio > 0)%>%
  ggplot(survey2016, mapping = aes(x = WaistHipratio)) + geom_histogram(binwidth = 0.005) + labs(x="WaistHipRatio",y="Count")
```
### bivariate assioation
```{r walking_whval_ggmarginal}
#Number of days walking 30 mins+ fast or brisk in last 4 weeks
hexplot1 <- survey2016 %>% 
  ggplot(aes(x=walking, y=WaistHipratio)) + 
  geom_point(position = "jitter") +
  labs(x="Walking",y="WaistHipratio")
  ggMarginal(hexplot1, type = "histogram")
```

```{r sport_whval_ggmarginal}
#Average minutes doing sports of 10 mins+ per week (vig mins*1)
hexplot2 <- survey2016 %>% 
  ggplot(aes(x=sport, y=WaistHipratio)) +  geom_point() +
  labs(x="Sport",y="WaistHipratio")
  ggMarginal(hexplot2, type = "histogram")
```

```{r exertion_whval_ggmarginal}
#Occasions 10+min muscle-strengthening in 4 weeks
hexplot3 <-survey2016 %>% 
  ggplot(aes(x=exertion, y=WaistHipratio)) +  geom_point(position = "jitter")+
  labs(x="exertion",y="WaistHipratio")
  ggMarginal(hexplot3, type = "histogram")

```

```{r sedentary_whval_ggmarginal}
#Total sedentary time on weekday (mins). As there is only 1440 minutes in one day, remove those data larger than 1440minutes.

hexplot4 <-survey2016 %>% 
  ggplot(aes(x=sedentary, y=WaistHipratio)) +  
  geom_point(position = "jitter")+
  labs(x="sedentary",y="WaistHipratio")

  ggMarginal(hexplot4, type = "histogram")

```

```{r drinkfrequency_whval_ggmarginal}
#Frequency drink alcohol in past 12 months(including non-drinkers) including 8 groups arranging from 1 to 8.The bigger of the figure, less frequency of drinking.
hexplot5 <-survey2016 %>% 
  ggplot(aes(x=drinkfrequency, y=WaistHipratio)) +  
  geom_point(position = "jitter")+
  labs(x="drinkfrequency",y="WaistHipratio")

  ggMarginal(hexplot5, type = "histogram")

```

```{r drinktotal_whval_ggmarginal}
#Total unit of alcohol per week.
hexplot6 <-survey2016 %>% 
  ggplot(aes(x=drinktotal, y=WaistHipratio)) +  
  geom_point(position = "jitter")+
  labs(x="drinktotal",y="WaistHipratio")

  ggMarginal(hexplot6, type = "histogram")
```

```{r fruitandveg_whval_ggmarginal}
#Total portion of fruit and veg.
hexplot7 <-survey2016 %>% 
  ggplot(aes(x=fruitandveg, y=WaistHipratio)) +  
  geom_point(position = "jitter")+
  labs(x="fruitandveg",y="WaistHipratio")

  ggMarginal(hexplot7, type = "histogram")
```

```{r all_whval_cor}
survey2016 %>%
  select(walking , sport, exertion, sedentary, drinkfrequency, drinktotal, fruitandveg, WaistHipratio ) %>% cor()
```
As is shown in the table, sedentary(Total sedentary time on weekday) and drinktotal(Total unit of alcohol per week) are positively correlated with the dependent variable, while walking(Number of days walking 30 mins+ fast or brisk in last 4 weeks), sport(Average minutes doing sports of 10 mins+ per week), exertion(Occasions 10+min muscle-strengthening in 4 weeks) and drinkfrequency(Frequency drink alcohol in past 12 months) are negatively correlated with the dependent variable.

Among these independent variables, the relationship between walking and dependent variables is beyond expectations. Now, check the monotonicity.

### hypothesis testing
```{r sex_whval}
survey2016 <- mutate(survey2016,
                     Sex = factor(Sex,
                                  levels=1:2,
                                  labels=c("Male","Female")))
survey2016 %>% 
  ggplot(aes(Sex, WaistHipratio)) + geom_boxplot()

survey2016 %>% t.test(WaistHipratio ~ Sex, data=.)
```
The difference of sex could influnce the differece of Waist/Hip ratio. 


```{r cigarette_whval}
#Cigarette Smoking Status
survey2016 <- mutate(survey2016,
                     cigarette = factor(cigarette,
                                  levels=1:5,
                                  labels=c("1","2","3","4","5")))


survey2016 %>% 
  ggplot(aes(cigarette,WaistHipratio)) + geom_boxplot() +
  labs(x = "cigarette", y = "WaistHipratio")

survey2016 %>% aov(WaistHipratio ~ cigarette, data=.) %>% summary
```

```{r age_whval}
#Age35g is respondent age grouped by approximately 3 year bands with 22 response categories. 7 16-19;8 20-24;9 25-29;...;22 90+ 
survey2016 <- mutate(survey2016,
                     age = factor(age,
                                  levels=7:22,
                                  labels=c("7","8","9","10","11","12","13","14","15","16","17","18","19","20","21","22")))


survey2016 %>% 
  ggplot(aes(age, WaistHipratio)) + geom_boxplot() +
  labs(x = "Age", y = "WaistHipratio")

survey2016 %>% aov(WaistHipratio ~ age, data=.) %>% summary
```
The difference of age of adults could influnce the differece of Waist/Hip ratio. 

```{r education_whval}
#highest education qualification
survey2016 <- mutate(survey2016,
                     education = factor(education,
                                  levels=1:7,
                                  labels=c("1","2","3","4","5","6","7")))


survey2016 %>% 
  ggplot(aes(education,WaistHipratio)) + geom_boxplot() +
  labs(x = "education", y = "WaistHipratio")

survey2016 %>% aov(WaistHipratio ~ education, data=.) %>% summary
```
The difference of education level could influnce the differece of Waist/Hip ratio.

```{r employment_whval}
#
survey2016 <- mutate(survey2016,
                     employment = factor(employment,
                                  levels=1:8,
                                  labels=c("1","2","3","4","5","6","7","8")))


survey2016 %>% 
  ggplot(aes(employment, WaistHipratio)) + geom_boxplot() +
  labs(x = "employment", y = "WaistHipratio")

survey2016 %>%
  aov(WaistHipratio ~ employment, data=.) %>% summary

```
The difference of employment status could influnce the differece of Waist/Hip ratio.

```{r income_whval}
survey2016 <- mutate(survey2016,
                     income = factor(income ,
                                  levels=1:5,
                                  labels=c("1","2","3","4","5")))


survey2016 %>% 
  ggplot(aes(income, WaistHipratio)) + geom_boxplot() +
  labs(x = "income", y = "WaistHipratio")

survey2016 %>%
  aov(WaistHipratio ~ income, data=.) %>% summary

```
The difference of income status could influnce the differece of Waist/Hip ratio.


```{r gambling_whval}
#Number of gambling activities participated in within last 12 months
survey2016 <- mutate(survey2016,
                     gambling = factor(gambling,
                                  levels=0:8,
                       
                                             labels=c("0","1","2","3","4","5","6","7","8")))

survey2016 %>% 
  ggplot(aes(gambling, WaistHipratio)) + geom_boxplot() +
  labs(x = "gambling", y = "WaistHipratio")

survey2016 %>%
  aov(WaistHipratio ~ gambling, data=.) %>% summary

```

###lm building

```{r model11}
survey2016 %>%
  lm(WaistHipratio ~ walking, data=.) -> model11
  summary(model11)
```

```{r model12}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport, data=.) -> model12
  summary(model12)
```


```{r model13}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion, data=.) -> model13
  summary(model13)
# Each coefficient is influenced by the other variables in a regression model.Because predictor variables are nearly always associated, two or more variables may explain some of the same variation in Y.
```

```{r model14}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary, data=.) -> model14
  summary(model14)
```
```{r model15}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency, data=.) -> model15
  summary(model15)
```
```{r model16}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal, data=.) -> model16
  summary(model16)
```
```{r model17}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex, data=.) -> model17
  summary(model17)
```

```{r model18}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex + age, data=.) -> model18
  summary(model18)
```

```{r model19}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex + age + education, data=.) -> model19
  summary(model19)
```
```{r model110}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex + age + education + employment, data=.) -> model110
  summary(model110)
```
```{r model111}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex + age + education + employment + income, data=.) -> model111
  summary(model111)
```
sex contributes more on the explaination of the variance of dependant variable compared with other independent variable, cause the adjusted R squre improves significantly after adding Sex in the model.

```{r model112}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex + age + education + employment + income + cigarette, data=.) -> model112
  summary(model112)
```

```{r model113}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex + age + education + employment + income + cigarette + gambling, data=.) -> model113
  summary(model113)
```

```{r model114}
survey2016 %>%
  lm(WaistHipratio ~ walking + sport + exertion +sedentary + drinkfrequency + drinktotal + Sex + age + education + employment + income + cigarette + gambling + fruitandveg, data=.) -> model114
  summary(model114)
```



```{r lm114log}
survey2016 %>%
  lm(log(WaistHipratio) ~ walking + sport + exertion + sedentary + drinkfrequency + drinktotal + Sex + age + education + employment + income + cigarette + fruitandveg + gambling, data=.) -> model114log
  summary(model114log)
```


###assumption testing
```{r specified_reset+fitted}
  resettest(model114)
  plot(model114)
#Residuals vs Fitted: This plot shows if residuals have non-linear patterns. There could be a non-linear relationship between predictor variables and an outcome variable and the pattern could show up in this plot if the model doesn’t capture the non-linear relationship.If you find equally spread residuals around a horizontal line without distinct patterns, that is a good indication you don’t have non-linear relationships.
#Normal Q-Q:This plot shows if residuals are normally distributed. Do residuals follow a straight line well or do they deviate severely? It’s good if residuals are lined well on the straight dashed line.
#Scale-Location: It’s also called Spread-Location plot. This plot shows if residuals are spread equally along the ranges of predictors. This is how you can check the assumption of equal variance (homoscedasticity). It’s good if you see a horizontal line with equally (randomly) spread points.
#Residuals vs Leverage:We watch out for outlying values at the upper right corner or at the lower right corner. Those spots are the places where cases can be influential against a regression line. Look for cases outside of a dashed line, Cook’s distance. When cases are outside of the Cook’s distance (meaning they have high Cook’s distance scores), the cases are influential to the regression results. The regression results will be altered if we exclude those cases.    
```

```{r homoscedasticity}
bptest(model114)
coeftest(model114, vcov = vcovHC(model114))
```


```{r independent}
hist(model114$residuals)
plot(model114, which = 2)
```

```{r multicollinearity}
vif(model114, digits = 3)
```



